DSSynth: An Automated Digital Controller Synthesis Tool for Physical Plants

### 2017-07-08 ase2017/paper.tex

In Step~$1$, the user selects the digital controller representation, which
can be a transfer function or a state-space model.  In Step~$2$, the
physical plant (in the form of Equation~(\ref{eq:transferfunction}) or
(\ref{eq:ss-example}), depending on representation) must be
given~\cite{astrom1997computer}.  Finally, in Step~$3$, the numerical
representation for the digital controller implementation must be set by the
user, i.e., the FWL format that defines the number of bits of the integer
and fractional parts as well as the dynamic range inputs.  The result of the
input given by the user is a specification of the physical plant.

After that, the automated synthesis process starts with Step~$A$, where the
\tool translates the digital system specification from the MATLAB format
into an ANSI-C program.  In Step~$B$, our CEGIS engine is invoked to
synthesize the digital controller w.r.t.~the digital system specification. 
Finally, in Step~$C$, the synthesized digital controller is produced.  The
output generated by the \tool is the synthesized digital controller
represented either in transfer function or state-space equation.  The
synthesis is considered to be \emph{successful} if a digital controller is
correctly synthesized w.r.t.~FWL effects; otherwise, if any parameter is
incorrectly defined in previous steps, or if there is no solution (no
controller) for the respective physical plant (given the implementation
requirements), then the synthesis is considered to have \emph{failed}.

Our CEGIS engine is implemented as an integrated module within the C bounded
model checker (CBMC)~\cite{cbmc}.  CBMC transforms the ANSI-C representation
of our closed-loop control system model into its internal format called
\emph{GOTO}.  We instrument this internal \emph{GOTO} program for each
synthesis or verification scenario accordingly, and use CBMC as an oracle to
answer our queries.  CBMC itself relies on an underlying SAT or SMT solver
to solve the resulting formulae.  We~model FWL effects, time discretization
and the presence of quantization noise explicitly using CBMC's
nondeterminism API (e.g., \texttt{nondet}, \texttt{CPROVER\_assume}
intrinsic functions).  Our module was included in CBMC~5.7 and is available
for public
download.\footnote{\url{https://github.com/diffblue/cbmc/archive/cbmc-5.7.zip}}
We further created a VirtualBox OVA image with our tool pre-installed,
including our MATLAB-generated benchmarks and benchmark shell script with
instructions to run all benchmarks and reproduce our
experiments.\footnote{\url{www.cprover.org/DSSynth/controller-synthesis-cav-2017.tar.gz}}
The current development branch of our toolbox is available in a public
GitHub
repository.\footnote{\url{https://github.com/ssvlab/dsverifier/tree/master/toolbox-dssynth}}

The entire implementation consists of $400$ lines of MATLAB scripts for the
UI module, $585$ and $800$ lines of C code in the naive and abstract
accelerator back-ends, respectively (cf.~Section~\ref{sec:CEGIS}).  $4000$
lines of C code for utility functions such as interval and fixed-point
arithmetic, as well as $1000$ lines of C++ code in the control synthesis
module in CEGIS CBMC.  We evaluate \tool using $18$ SISO control system
benchmarks.  Additionally, the documentation and videos about \tool are
available at \url{http://dsverifier.org/dsverifier/dssynth-toolbox/}.

@@

The \tool performs the following automated procedures 
to synthesize a given digital controller:

\begin{enumerate}
\item \textbf{Setup}: obtains the physical plant, fixed-point format 
and dynamic input ranges, and translates them to a specific structure in MATLAB.
\item \textbf{Parse}: obtains the digital system specification and translates 
it into an ANSI-C program.
\item \textbf{Execution}: obtains the ANSI-C program from the previous (parse) step 
and calls our CEGIS engine as a back-end program synthesis tool to perform the automated synthesis;
\item \textbf{Extraction}: obtains the ``.log'' file that is generated 
after the synthesis phase and then checks the synthesized digital-controller.
\item \textbf{Report}: obtains the digital controller from the previous (extraction) step 
and translates it into the MATLAB structures to show the respective result to the user.
\end{enumerate}

@@

The features of the \tool can be summarized as follows:

\begin{itemize}
\item \textbf{Representation}: the synthesis can be 
performed for digital systems specified via 
transfer function or as state-space models.
\item \textbf{Fixed-Point Arithmetic}: the digital 
controller is synthesized considering FWL effects, 
using the given fixed-point numerical representation.
\item \textbf{Interface}: the synthesis can be executed 
via command-line or graphical-user interface (GUI). 
For the GUI application, the user can plot the step response 
to reproduce the stability in closed-loop of the synthesized digital controller.
\end{itemize}

@@

Users must provide a digital system described as a MATLAB model 
using a \texttt{tf} (for transfer function) or an \texttt{ss} (for state-space model) 
command (cf. step $1$ of Fig.~\ref{fig:synthesis-flow}).
\tool is called via command line in MATLAB as

@@

\noindent where \texttt{plant} is the physical plant in state-space or transfer function representation, 
\texttt{intBits} is the integer part, \texttt{fracBits} is the fractional part, \texttt{maxR} and \texttt{minR} 
are the maximum and minimum dynamic range, respectively (cf. steps $2$ and $3$ of Fig.~\ref{fig:synthesis-flow}).
%
After executing the \texttt{synthesize} command, 
\tool provides the synthesized digital controller, according to the implementation aspects.

@@

A graphical user interface application was developed (illustrated in
Fig.~\ref{fig:gui-for-tf}) to favor digital controller synthesis in MATLAB. 
Its aim is to improve usability and, consequently, to attract more digital
dynamical system engineers.  Users can provide all required parameters for
digital controller synthesis: physical plant specification, fixed-point
representation and dynamic inputs.
%
\begin{figure}[ht]
    \centering
    \subfloat[Define the system representation and the physical plant.]{
		\includegraphics[width=0.2\textwidth]{step1.png}
		\label{step1}}
		\hfil
    \subfloat[Define implementation aspects and dynamic range.]{
	\includegraphics[width=0.2\textwidth]{step2.png}
		\label{step2}}
		\hfil
    \subfloat[Resulting digital controller synthesized by DSSynth.]{
	\includegraphics[width=0.2\textwidth]{step3.png}
		\label{step2}}
		\hfil
    \subfloat[Step response for the synthesized digital controller.]{
	\includegraphics[width=0.22\textwidth]{step4.png}
		\label{step3}}
		\hfil
    \caption{DSSynth GUI application for the controller synthesis based on transfer function representation.}
    \label{fig:gui-for-tf}
\end{figure}

@@

To illustrate the \tool's usage, Fig.~\ref{toolbox-usage} shows each step of the 
synthesis process for a physical plant described in Eq.~\eqref{equation_plant},
which represents a real unmanned aerial vehicle (UAV) quadcopter system~\cite{bouabdallah}. 
This physical plant is expressed by a transfer function $G(z)$, where 
$B(z)$ and $A(z)$ represent the numerator and denominator, respectively, as
%
\begin{equation}
\label{equation_plant}
G(z)=\frac{B(z)}{A(z)}=\frac{-0.06875z^{2}}{z^2-1.696z+0.7089}.
\end{equation}

\begin{figure}[ht]
\scriptsize
\begin{lstlisting}[xleftmargin=.025\textwidth,xrightmargin=.025\textwidth, frame=single,]
>> num = [-0.06875 0 0];
>> den = [1.0000 -1.696 0.7089];
>> system = tf(num,den,0.002);
>> y = synthesize(system,8,8,1,-1);
>> SYNTHESIS SUCCESSFUL
>> y = 
>>  -0.9983z^2 + 0.09587z + 0.1926
>>  --------------------------------
>>       z^2 + 0.5665z + 0.75
\end{lstlisting}
\vspace{-0.2cm}
\caption{Synthesis of a digital controller for the physical plant defined in Eq.(\ref{equation_plant}) in MATLAB, with a fixed-point format  $\left\langle 8,8\right\rangle$.}
\label{toolbox-usage}
\end{figure}

@@

A digital system is stable \textit{iff} all of its poles are inside the
$z$-plane unity circle, namely poles must have the modulus less than
one~\cite{astrom1997computer}.  In order to compute the poles for the
digital system represented by Equations~\eqref{equation_plant} and
\eqref{equation_controller}, the user can invoke the \texttt{feedback} and
\texttt{series} commands in MATLAB~\cite{matlab-toolbox} to achieve a
general equation regarding the transfer function representation, as
illustrated in Fig.~\ref{combine-controller-plant}.

@@

Simulating the digital system represented by Equations
\eqref{equation_plant} and \eqref{equation_controller} to validate and
reproduce the system stability, the user can invoke the step response using
the command \texttt{dstep} in MATLAB, and then observe that the plotted
graph shows a stable digital system, as illustrated in
Fig.~\ref{step-response}.
%
\begin{figure}[ht]
  \includegraphics[width=0.5\textwidth]{step-response.eps}
  \caption{Step response for the Eq.~\eqref{equation_plant} describing a stable system for the UAV quadcopter system.}
  \label{step-response}
\end{figure}

@@

Our evaluation consists of $18$ SISO control system benchmarks extracted
from the literature~\cite{abate2017,abatecav2017,bouabdallah,
acrobot,cstr,KOKOTOVIC198023,gajic2008optimal,Franklin15, maglev,
converters,
CTMS,DBLP:journals/tc/BessaIPCF17,DBLP:journals/dafes/BessaICF16}, which
include physical plants for unmanned aerial vehicle (UAV), cruise control
system, spring-mass damper, satellite, direct current (DC) servo motor,
helicopter, inverted pendulum, pendulum, magnetic suspension, magnitezed
pointer, 1/4 car suspension, computer tape driver, flexible beam, guidance
system, USCG cutter tampa, continuous stirred tank reactor, voltage
regulator, and an acrobot control system.

@@

For all evaluated benchmarks, the signal input range related to the control
system lies between $-1$ and $1$.  The implementation features used for
digital controller are: $8$-bits for the integer part (including the sign
bit) and $8$-bits for the fractional part.  We have run the \tool using the
CEGIS with multi-staged verification (cf.~Section~\ref{sec:CEGIS}).  All
experiments with \tool v$1$.$0$.$0$ were conducted on an otherwise idle
Intel Core i$7-2600$ $3.40$ GHz processor, with $24$ GB of RAM, running
Ubuntu $64$-bits.

@@

The main objectives of our experimental evaluation is to (1) evaluate the \tool performance 
to produce stable and safe controllers, and also (2) confirm the stability and safety of the synthesized digital 
controllers outside of our model using MATLAB. %, using both the transfer function and state-space representations.

>>>>

In Step~1, the user selects the digital controller representation, which
can be a transfer function or a state-space model.  In Step~2, the
physical plant (in the form of Equation~(\ref{eq:transferfunction}) or
(\ref{eq:ss-example}), depending on representation) must be
given~\cite{astrom1997computer}.  Finally, in Step~3, the numerical
representation for the digital controller implementation must be set by the
user, i.e., the FWL format that defines the number of bits of the integer
and fractional parts as well as the dynamic range inputs.  The result of the
input given by the user is a specification of the physical plant.

After that, the automated synthesis process starts with Step~A, where the
\tool translates the digital system specification from the MATLAB format
into an ANSI-C program.  In Step~B, our CEGIS engine is invoked to
synthesize the digital controller w.r.t.~the digital system specification. 
Finally, in Step~C, the synthesized digital controller is produced.  The
output generated by the \tool is the synthesized digital controller
represented either in transfer function or state-space equation.  The
synthesis is considered to be \emph{successful} if a digital controller is
correctly synthesized w.r.t.~FWL effects; otherwise, if any parameter is
incorrectly defined in previous steps, or if there is no solution (no
controller) for the respective physical plant (given the implementation
requirements), then the synthesis is considered to have \emph{failed}.

Our CEGIS engine is implemented as an integrated module within the C bounded
model checker (CBMC)~\cite{cbmc}.  CBMC transforms the ANSI-C representation
of our closed-loop control system model into its internal format.  We
instrument this IR for each synthesis or verification scenario accordingly,
and use CBMC as an oracle to answer our queries.  CBMC itself relies on an
underlying SAT or SMT solver to solve verification conditions.  We~model FWL
effects, time discretization and the presence of quantization noise
explicitly using CBMC's nondeterminism API (e.g., \texttt{nondet},
\texttt{CPROVER\_assume} intrinsic functions).  Our module is included in
CBMC~5.7 and is available for
download.\footnote{\url{https://github.com/diffblue/cbmc/archive/cbmc-5.7.zip}}
We further created a VirtualBox OVA image with our tool pre-installed,
including our MATLAB-generated benchmarks and benchmark shell script with
instructions to run all benchmarks and reproduce our
experiments.\footnote{\url{www.cprover.org/DSSynth/controller-synthesis-cav-2017.tar.gz}}
The current development branch of our toolbox is available in a public
GitHub
repository.\footnote{\url{https://github.com/ssvlab/dsverifier/tree/master/toolbox-dssynth}}

The entire implementation consists of $400$ lines of MATLAB scripts for the
UI module, $585$ and $800$ lines of C code in the naive and abstract
accelerator back-ends, respectively (cf.~Section~\ref{sec:CEGIS}), $4000$
lines of C code for utility functions such as interval and fixed-point
arithmetic, as well as $1000$ lines of C++ code in the control synthesis
module in CEGIS CBMC.  We evaluate \tool using $18$ SISO control system
benchmarks.  Additionally, documentation and videos about the \tool are
available at \url{http://dsverifier.org/dsverifier/dssynth-toolbox/}.

@@

The \tool performs the following automated steps
to synthesize a given digital controller:
%
\begin{enumerate}
%
\item \textbf{Setup}: obtains the physical plant, fixed-point format and
dynamic input ranges, and translates them to a specific structure in MATLAB.
%
\item \textbf{Parse}: obtains the digital system specification and translates 
it into an ANSI-C program.
%
\item \textbf{Execution}: obtains the ANSI-C program from the previous
(parse) step and calls our CEGIS engine as a back-end program synthesis tool
to perform the automated synthesis;
%
\item \textbf{Extraction}: obtains the ``.log'' file that is generated 
after the synthesis phase and then checks the synthesized digital-controller.
%
\item \textbf{Report}: obtains the digital controller from the previous
(extraction) step and translates it into MATLAB, for presentation to the
user.
%
\end{enumerate}

@@

The features of the \tool can be summarized as follows:
%
\begin{itemize}
%
\item \textbf{Representation}: the synthesis can be 
performed for digital systems specified via 
transfer function or as state-space models.
%
\item \textbf{Fixed-Point Arithmetic}: the digital 
controller is synthesized considering FWL effects, 
using the given fixed-point numerical representation.
%
\item \textbf{Interface}: the synthesis can be executed via command-line or
graphical-user interface (GUI).  For the GUI application, the user can plot
the step response to reproduce the stability in closed-loop of the
synthesized digital controller.
%
\end{itemize}

@@

Users must provide a digital system described as a MATLAB model using a
\texttt{tf} (for transfer function) or an \texttt{ss} (for state-space
model) command (step $1$ of Fig.~\ref{fig:synthesis-flow}).  The \tool is
called via the command line in MATLAB as

@@

\noindent where \texttt{plant} is the physical plant in state-space or
transfer function representation, \texttt{intBits} is the integer part,
\texttt{fracBits} is the fractional part, \texttt{maxR} and \texttt{minR}
are the maximum and minimum dynamic range, respectively (steps $2$ and
$3$ of Fig.~\ref{fig:synthesis-flow}).
%
After executing the \texttt{synthesize} command, the \tool provides the
synthesized digital controller.

@@

A graphical user interface application was developed (illustrated in
Fig.~\ref{fig:gui-for-tf}) to support digital controller synthesis in
MATLAB.  Its aim is to improve usability and, consequently, to attract more
engineers.  Users provide the following inputs for digital controller
synthesis within the UI: physical plant specification, fixed-point
representation and the range of the inputs.

\begin{figure}[ht]
    \centering
    \subfloat[Definition of the system representation and the physical plant]{
		\includegraphics[width=0.2\textwidth]{step1.png}
		\label{step1}}
		\hfil
    \subfloat[Definition of implementation aspects and input ranges]{
	\includegraphics[width=0.2\textwidth]{step2.png}
		\label{step2}}
		\hfil
    \subfloat[Digital controller synthesized by DSSynth]{
	\includegraphics[width=0.2\textwidth]{step3.png}
		\label{step2}}
		\hfil
    \subfloat[Step response for the synthesized digital controller]{
	\includegraphics[width=0.22\textwidth]{step4.png}
		\label{step3}}
		\hfil
    \caption{DSSynth GUI for controller synthesis using transfer function representation}
    \label{fig:gui-for-tf}
\end{figure}

@@

To illustrate the usage of the \tool, Fig.~\ref{toolbox-usage} shows each
step of the synthesis process for a physical plant described in
Eq.~\eqref{equation_plant}, which represents a unmanned aerial vehicle
(UAV) quadcopter system~\cite{bouabdallah}.  This physical plant is
expressed by a transfer function $G(z)$, where $B(z)$ and $A(z)$ represent
the numerator and denominator, respectively, as
%
\begin{equation}
\label{equation_plant}
G(z)=\frac{B(z)}{A(z)}=\frac{-0.06875z^{2}}{z^2-1.696z+0.7089}.
\end{equation}

\begin{figure}[ht]
\scriptsize
\begin{lstlisting}[xleftmargin=.025\textwidth,xrightmargin=.025\textwidth, frame=single,]
>> num = [-0.06875 0 0];
>> den = [1.0000 -1.696 0.7089];
>> system = tf(num,den,0.002);
>> y = synthesize(system,8,8,1,-1);
>> SYNTHESIS SUCCESSFUL
>> y = 
>>  -0.9983z^2 + 0.09587z + 0.1926
>>  --------------------------------
>>       z^2 + 0.5665z + 0.75
\end{lstlisting}
\vspace{-0.2cm}
\caption{Synthesis of a digital controller for the physical plant defined in Eq.(\ref{equation_plant}) in MATLAB, with a fixed-point format  $\left\langle 8,8\right\rangle$}
\label{toolbox-usage}
\end{figure}

@@

A digital system is stable iff all of its poles are inside the
$z$-plane unity circle, i.e., the poles must have a modulus less than
one~\cite{astrom1997computer}.  In order to compute the poles for the
digital system represented by Equations~\eqref{equation_plant} and
\eqref{equation_controller}, the user can invoke the \texttt{feedback} and
\texttt{series} commands in MATLAB~\cite{matlab-toolbox} to obtain a
general equation regarding the transfer function representation, as
illustrated in Fig.~\ref{combine-controller-plant}.

@@

Simulating the digital system represented by Equations
\eqref{equation_plant} and \eqref{equation_controller} to validate and
reproduce the system stability, the user can obtain the step response using
the command \texttt{dstep} in MATLAB, and then observe that the plotted
graph indicates a stable digital system (Fig.~\ref{step-response}).

\begin{figure}[ht]
  \includegraphics[width=0.5\textwidth]{step-response.eps}
  \caption{Step response for Eq.~\eqref{equation_plant} describing a stable system for the UAV quadcopter system}
  \label{step-response}
\end{figure}

@@

Our evaluation consists of $18$ SISO control system benchmarks extracted
from the literature~\cite{abate2017, abatecav2017, bouabdallah, acrobot,
cstr, KOKOTOVIC198023, gajic2008optimal, Franklin15, maglev, converters,
CTMS, DBLP:journals/tc/BessaIPCF17, DBLP:journals/dafes/BessaICF16}, which
include physical plants for a unmanned aerial vehicle (UAV), a cruise
control system, a spring-mass damper, a satellite, a direct current (DC)
servo motor, a helicopter, the inverted pendulum, a pendulum, magnetic
suspension, a magnitezed pointer, 1/4 car suspension, a computer tape
driver, a flexible beam, a guidance system, a USCG cutter tampa, a
continuous stirred tank reactor, a voltage regulator, and an acrobot control
system.

@@

For all evaluated benchmarks, the signal input range related to the control
system is between $-1$ and $1$.  The implementation features used for
digital controller are: $8$~bits for the integer part (including the sign
bit) and $8$~bits for the fractional part.  We have run the \tool using
CEGIS with multi-staged verification (cf.~Section~\ref{sec:CEGIS}).  All
experiments with \tool v$1$.$0$.$0$ were conducted on an otherwise idle
Intel Core i$7-2600$ $3.40$ GHz processor, with $24$ GB of RAM, running
Ubuntu $64$-bits.

@@

The main objectives of our experimental evaluation is to (1)~evaluate the
\tool performance to produce stable and safe controllers, and also
(2)~confirm the stability and safety of the synthesized digital controllers
outside of our model using MATLAB.

%, using both the transfer function and state-space representations.
